---
title: "House_Prices"
author: "Angelos Theodorakis"
date: "29/5/2020"
output: 
  md_document:
          variant: markdown_github
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=FALSE,message=FALSE)
```

First, we will load the libraries we need.

```{r}
library(plyr)
library(dplyr)
```

Then we load the data, the train and the test set seperately.

```{r}
setwd("C:/Users/User/Desktop/Άγγελος/R/Data analysis/House prices/house-prices-advanced-regression-techniques")
train<-read.csv("train.csv",stringsAsFactors = FALSE)
test<-read.csv("test.csv",header=TRUE,stringsAsFactors = FALSE)
```

We can add the SalePrice variable in the test set as NA.

```{r}
test$SalePrice <- NA
```

Now we combine the test and train set and explore our data.

```{r}
data <- rbind(train,test)
str(data)
```

Let's explore SalePrice variable, which is the one we want to predict.

```{r}
summary(data$SalePrice) # We can see that the median for the sales price is 163000
data$SalePrice<-as.numeric(data$SalePrice) # Change to numeric
hist(data$SalePrice, breaks=30, xlab = 'Price', main = 'Histogram of Sale Price') # There is a slightly skewed distribution to the right, possibly because of the extreme values.
```

We will now check for missing values.

```{r}
sort(sapply(data, function(x) sum(is.na(x))),decreasing=TRUE)
plot(sort(sapply(data, function(x) sum(is.na(x))),decreasing=TRUE),type='h')
```

There are some columns with lots of missing values. However, in most cases these are not missing values, but an indication that the apartment doesn't have these amenities. So let's replace these Na's with 'None' and explore these variables.

**PoolQC:  Pool quality**
<br>
This is the first variable we will examine. From the documentation, we observe that indeed, the 'NA' value in this variable means that there is no pool. We will also replace the values with integers from 0 to 5 as long as it is a scaling variable.<br>
We will continue in a similar way with the other variables that have missing values.

```{r}
data$PoolQC[is.na(data$PoolQC)] <- "None"
table(data$PoolQC)
Qual_Cond <- c('None' = 0, 'Po' = 1, 'Fa' = 2, 'TA' = 3, 'Gd' = 4, 'Ex' = 5)
data$PoolQC<-revalue(data$PoolQC,Qual_Cond)
data$PoolQC<-as.integer(data$PoolQC)
```

**MiscFeature: Miscellaneous feature not covered in other categories**

```{r}
table(data$MiscFeature)
data$MiscFeature[is.na(data$MiscFeature)] <- "None"
table(data$MiscFeature)
data$MiscFeature<-as.factor(data$MiscFeature) 
plot(data$MiscFeature)
```

**Alley: Type of alley access to property** 

```{r}
table(data$Alley)
data$Alley[is.na(data$Alley)] <- "None"
data$Alley<-as.factor(data$Alley)
plot(data$Alley,data$SalePrice) #Seems to affect the price
table(data$Alley)
```

**Fence: Fence quality**

```{r}
table(data$Fence)
data$Fence[is.na(data$Fence)] <- "None"
data$Fence<-as.factor(data$Fence)
plot(data$Fence,data$SalePrice) # Seems to affect the price. The 'no fence' variable has the highest median!
table(data$Fence)
```

**FireplaceQu: Fireplace quality**

```{r}
table(data$FireplaceQu)
data$FireplaceQu[is.na(data$FireplaceQu)] <- "None"
data$FireplaceQu<-revalue(data$FireplaceQu,Qual_Cond) #replace the values with Quality/Condition values
data$FireplaceQu<-as.integer(data$FireplaceQu)
plot(data$FireplaceQu,data$SalePrice) #Seems to affect the price!
abline(lm(data$SalePrice~data$FireplaceQu))
table(data$FireplaceQu)
```

**LotFrontage: Linear feet of street connected to property.**

 There are 259 NA's. We will replace the missing values with the median and not the mean, because of the extreme values (outliers).

```{r}
summary(data$LotFrontage)
cor(data$LotFrontage,data$SalePrice, use = "pairwise.complete.obs")
hist(data$LotFrontage,breaks=30,xlim = c(0,200))
plot(data$LotFrontage,data$SalePrice) # There are 2 extreme values
data$LotFrontage[is.na(data$LotFrontage)]<-median(na.omit(as.numeric(data$LotFrontage)))
```

**GarageYrBlt**

```{r}
summary(data$GarageYrBlt)
plot(data$GarageYrBlt) 
```

There is a wrong observation unless it is a house from the future (2207)! We replace it with the year 2007.

```{r}
which(data[,"GarageYrBlt"]>2019)
data[2593,"GarageYrBlt"]<-2007 # Replace the observation
plot(as.factor(data$GarageYrBlt),data$SalePrice,xlab = 'Year the garage was built', ylab = 'House Price')
```

We will find out the NA's by examining another variable, YearBuilt. Let's see if they have a correlation.

```{r}
head(data.frame(data$YearBuilt,data$GarageYrBlt),15)
length(which(data$YearBuilt!=data$GarageYrBlt)) # Number of different rows
```

Indeed the year the house was built is in most cases the same as the year the garage was built, so we will drop the variable and keep the YearBuilt variable.

```{r}
data <- subset(data, select = -GarageYrBlt)
```

**GarageFinish:Interior finish of the garage**

```{r}
plot(as.factor(data$GarageFinish))
plot(as.factor(data$GarageFinish),data$SalePrice)
data$GarageFinish[is.na(data$GarageFinish)] <- "None"
plot(as.factor(data$GarageFinish),data$SalePrice) # It seems ordinal 
levels(as.factor(data$GarageFinish))
Finish <- c('None' = 0, "Unf" = 1, "RFn" = 2, "Fin" = 3)
data$GarageFinish<-revalue(data$GarageFinish,Finish)
data$GarageFinish<-as.integer(data$GarageFinish)  
plot(data$GarageFinish,data$SalePrice)
```

**GarageQual: Garage quality**

```{r}
table(data$GarageQual)
plot(as.factor(data$GarageQual),data$SalePrice)
data$GarageQual[is.na(data$GarageQual)] <- "None"
plot(as.factor(data$GarageQual),data$SalePrice)
data$GarageQual<-revalue(data$GarageQual,Qual_Cond)
table(data$GarageQual)
data$GarageQual<-as.integer(data$GarageQual)
```

**GarageCond: Garage condition**

```{r}
table(data$GarageCond)
plot(as.factor(data$GarageCond),data$SalePrice)
data$GarageCond[is.na(data$GarageCond)] <- "None"
plot(as.factor(data$GarageCond),data$SalePrice)
data$GarageCond<-revalue(data$GarageCond,Qual_Cond)
table(data$GarageCond)
data$GarageCond<-as.integer(data$GarageCond)
```

We can see the correlation between garage condition and quality.

```{r}
cor(data$GarageCond,data$GarageQual)
```

It is a huge correlation,we must drop one of 2 variables

```{r}
cor(data[1:1459,'SalePrice'],data[1:1459,'GarageCond'])
cor(data[1:1459,'SalePrice'],data[1:1459,'GarageQual'])
```

We'll drop the variable less correlated with sales price

```{r}
data <- subset(data, select = -GarageCond)
```

**GarageType: Garage location**

```{r}
table(data$GarageType)
plot(as.factor(data$GarageType),data$SalePrice)
data$GarageType[is.na(data$GarageType)] <- "None"
plot(as.factor(data$GarageType),data$SalePrice)
data$GarageType<-as.factor(data$GarageType)
plot(data$GarageType,data$SalePrice)
```


**BsmtCond:Evaluates the general condition of the basement**

```{r}
table(data$BsmtCond)
plot(as.factor(data$BsmtCond),data$SalePrice)
data$BsmtCond[is.na(data$BsmtCond)] <- "None"
plot(as.factor(data$BsmtCond),data$SalePrice)
data$BsmtCond<-revalue(data$BsmtCond,Qual_Cond)
table(data$BsmtCond)
data$BsmtCond<-as.integer(data$BsmtCond)
plot(as.factor(data$BsmtCond),data$SalePrice)
```
There seems to be a correlation with the house price.

**BsmtExposure:Refers to walkout or garden level walls**

```{r}
table(data$BsmtExposure)
plot(as.factor(data$BsmtExposure),data$SalePrice)
data$BsmtExposure[is.na(data$BsmtExposure)] <- "None"
plot(as.factor(data$BsmtExposure),data$SalePrice)
Exposure<-c("None"=0,"No"=1,'Mn'=2,'Av'=3,'Gd'=4)
data$BsmtExposure<-revalue(data$BsmtExposure,Exposure)
table(data$BsmtExposure)
data$BsmtExposure<-as.integer(data$BsmtExposure)
plot(as.factor(data$BsmtExposure),data$SalePrice)
```



**BsmtQual:Evaluates the height of the basement**
```{r}
table(data$BsmtQual)
plot(as.factor(data$BsmtQual),data$SalePrice)
data$BsmtQual[is.na(data$BsmtQual)] <- "None"
plot(as.factor(data$BsmtQual),data$SalePrice)
data$BsmtQual<-revalue(data$BsmtQual,Qual_Cond)
table(data$BsmtQual)
data$BsmtQual<-as.integer(data$BsmtQual)
plot(as.factor(data$BsmtQual),data$SalePrice)
```

**BsmtFinType1:Rating of basement finished area**

```{r}
table(data$BsmtFinType1)
plot(as.factor(data$BsmtFinType1),data$SalePrice)
data$BsmtFinType1[is.na(data$BsmtFinType1)] <- "None"
plot(as.factor(data$BsmtFinType1),data$SalePrice)
Bsm_type<-c("None"=0,"Unf"=1,'LwQ'=2,'Rec'=3,'BLQ'=4,'ALQ'=5,'GLQ'=6)
data$BsmtFinType1<-revalue(data$BsmtFinType1,Bsm_type)
table(data$BsmtFinType1)
data$BsmtFinType1<-as.integer(data$BsmtFinType1)
plot(as.factor(data$BsmtFinType1),data$SalePrice)
```


**BsmtFinType2:Rating of basement finished area (if multiple types)**

```{r}
table(data$BsmtFinType2)
plot(as.factor(data$BsmtFinType2),data$SalePrice)
data$BsmtFinType2[is.na(data$BsmtFinType2)] <- "None"
plot(as.factor(data$BsmtFinType2),data$SalePrice)
data$BsmtFinType2<-revalue(data$BsmtFinType2,Bsm_type)
table(data$BsmtFinType2)
data$BsmtFinType2<-as.integer(data$BsmtFinType2)
plot(as.factor(data$BsmtFinType2),data$SalePrice)
```

**MasVnrType: Masonry veneer type (walls).**

```{r}
table(data$MasVnrType)
plot(as.factor(data$MasVnrType),data$SalePrice)
data$MasVnrType[is.na(data$MasVnrType)] <- "None" #Replace all Na's with 'none'
Masonry <- c('None'=0, 'BrkCmn'=1, 'BrkFace'=2, 'Stone'=3)
data$MasVnrType<-revalue(data$MasVnrType,Masonry)
table(data$MasVnrType)
data$MasVnrType<-as.integer(data$MasVnrType)
plot(as.factor(data$MasVnrType),data$SalePrice)
```

**MasVnrArea: Masonry veneer area in square feet**

```{r}
plot(data$MasVnrArea,data$SalePrice)
data$MasVnrArea[is.na(data$MasVnrArea)] <- 0
data$MasVnrArea<-as.numeric(data$MasVnrArea)
plot(data$MasVnrArea,data$SalePrice)
```

**MSZoning: Identifies the general zoning classification of the sale.**

```{r}
table(data$MSZoning)
plot(as.factor(data$MSZoning),data$SalePrice)
```

How we can find the missing values? Let's check out the MSSubClass

```{r}
data[is.na(data$MSZoning),c("MSZoning",'MSSubClass')]
table(data[,c("MSZoning",'MSSubClass')])
```

And now the correlations with numeric variables.

```{r}
library(corrplot)
cor(data$MSSubClass,data[,sapply(data, is.numeric)]) #
```

It is not higly correlated with other numeric variables.<br>
So by examining the above table, we will replace MSSubClass=20 with RL , MSSubClass=70 and MSSubClass=30 with RM.

```{r}
data$MSZoning[is.na(data$MSZoning)]<-c('RM','RL','RM','RL')
data[is.na(data$MSZoning),c("MSZoning",'MSSubClass')]
data$MSZoning<-as.factor(data$MSZoning)
```

**Utilities: Type of utilities available**

We don't need this variable for prediction , as there is only one house "NoSeWa".

```{r}
table(data$Utilities)
data <- subset(data, select = -Utilities)
```

**BsmtFullBath:Basement full bathrooms**


```{r}
table(data$BsmtFullBath)
```

Let's find all the basement variables below.

```{r}
paste(colnames(select(data,contains("Bsmt"))),collapse="','")
```


```{r}
data[is.na(data$BsmtFullBath),c('BsmtQual','BsmtCond','BsmtExposure','BsmtFinType1','BsmtFinSF1','BsmtFinType2','BsmtFinSF2','BsmtUnfSF','TotalBsmtSF','BsmtFullBath','BsmtHalfBath')]
```

We conclude that these variables are higly corralated and so we will replace the Na's with zero value


```{r}
data$BsmtFullBath[is.na(data$BsmtFullBath)]<-0 # Replace missing value with zero
plot(as.factor(data$BsmtFullBath),data$SalePrice)
data$BsmtFullBath<-as.integer(data$BsmtFullBath)
```

**BsmtHalfBath: Basement half bathrooms**

We proceed with the same way as above.

```{r}
table(data$BsmtHalfBath)
paste(colnames(select(data,contains("Bsmt"))),collapse="','")
data[is.na(data$BsmtHalfBath),c('BsmtQual','BsmtCond','BsmtExposure','BsmtFinType1','BsmtFinSF1','BsmtFinType2','BsmtFinSF2','BsmtUnfSF','TotalBsmtSF','BsmtFullBath','BsmtHalfBath')]
data$BsmtHalfBath[is.na(data$BsmtHalfBath)]<-0 # Replace missing value with zero
plot(as.factor(data$BsmtHalfBath),data$SalePrice)
data$BsmtHalfBath<-as.integer(data$BsmtHalfBath)
#we can see the correlation 
cor(data$BsmtHalfBath,data$BsmtFullBath) #It is not correlated
```

**Functional: Home functionality (Assume typical unless deductions are warranted)**

```{r}
table(data$Functional)
plot(as.factor(data$Functional),data$SalePrice)
```

We will replace the Na's with the most common value 7, which is typical functionality

```{r}
data$Functional[is.na(data$Functional)]<-7
Functional<-c('Sal'=0, 'Sev'=1, 'Maj2'=2, 'Maj1'=3, 'Mod'=4, 'Min2'=5, 'Min1'=6, 'Typ'=7)
data$Functional <- as.integer(revalue(data$Functional, Functional))
table(data$Functional)
```

**GarageCars Size of garage in car capacity** 

```{r}
table(as.factor(data$GarageCars))
plot(as.factor(data$GarageCars))
plot(as.factor(data$GarageCars),data$SalePrice)
paste(colnames(select(data,contains("Garage"))),collapse="','")
data[is.na(data$GarageCars),c('GarageType','GarageFinish','GarageCars','GarageArea','GarageQual')]
data$GarageCars[is.na(data$GarageCars)]<-0 # Replace missing value with zero
data$GarageCars<-as.integer(data$GarageCars)
```


**GarageArea**

```{r}
paste(colnames(select(data,contains("Garage"))),collapse="','")
data[is.na(data$GarageArea),c('GarageType','GarageFinish','GarageCars','GarageArea','GarageQual')]
data$GarageArea[is.na(data$GarageArea)]<-0 # Replace missing value with zero
data$GarageArea<-as.integer(data$GarageArea)
plot(data$GarageArea,data$SalePrice)
```

Let's find its correlation with other numeric values

```{r}
cor(data[,sapply(data, is.numeric)],data$GarageArea)
```

It is correlated with GarageCars (Rsq = 0.889) <br>
Should we drop GarageArea or GarageCars? <br>
We will examine which is less corellated with Sales Price.

```{r}
cor(data[1:1459,'SalePrice'],data[1:1459,'GarageArea'])
cor(data[1:1459,'SalePrice'],data[1:1459,'GarageCars']) 
```

We'll drop the GarageArea variable, since it is less corellated with sales Price 

```{r}
data <- subset(data, select = -GarageArea)
```

**TotalBsmtSF : Total square feet of basement area**

```{r}
plot(data$TotalBsmtSF,data$SalePrice)
cor(data[1:1459,'SalePrice'],data[1:1459,'TotalBsmtSF'])
data$TotalBsmtSF[is.na(data$TotalBsmtSF)]
paste(colnames(select(data,contains("Bsmt"))),collapse="','")
data[is.na(data$TotalBsmtSF),c('BsmtQual','BsmtCond','BsmtExposure','BsmtFinType1','BsmtFinSF1','BsmtFinType2','BsmtFinSF2','BsmtUnfSF','TotalBsmtSF','BsmtFullBath','BsmtHalfBath')]
data$TotalBsmtSF[is.na(data$TotalBsmtSF)]<-0 # Replace missing value with zero
```

**BsmtFinSF1: Type 1 finished square feet**   

```{r}
plot(data$BsmtFinSF1,data$SalePrice)
cor(data[1:1459,'SalePrice'],data[1:1459,'BsmtFinSF1'])
data$BsmtFinSF1[is.na(data$BsmtFinSF1)]
paste(colnames(select(data,contains("Bsmt"))),collapse="','")
data[is.na(data$BsmtFinSF1),c('BsmtQual','BsmtCond','BsmtExposure','BsmtFinType1','BsmtFinSF1','BsmtFinType2','BsmtFinSF2','BsmtUnfSF','TotalBsmtSF','BsmtFullBath','BsmtHalfBath')]
data$BsmtFinSF1[is.na(data$BsmtFinSF1)]<-0 # Replace missing value with zero
```

**BsmtUnfSF: Unfinished square feet of basement area**

```{r}
plot(data$BsmtUnfSF,data$SalePrice)
cor(data[1:1459,'SalePrice'],data[1:1459,'BsmtUnfSF'])
data$BsmtUnfSF[is.na(data$BsmtUnfSF)]
paste(colnames(select(data,contains("Bsmt"))),collapse="','")
data[is.na(data$BsmtUnfSF),c('BsmtQual','BsmtCond','BsmtExposure','BsmtFinType1','BsmtFinSF1','BsmtFinType2','BsmtFinSF2','BsmtUnfSF','TotalBsmtSF','BsmtFullBath','BsmtHalfBath')]
data$BsmtUnfSF[is.na(data$BsmtUnfSF)]<-0 # Replace missing value with zero
```

Let's see if it is correlated with another variable

```{r}
cor(data$BsmtUnfSF,data[,sapply(data, is.numeric)])
```

It is correlated with BsmtFinSF1 (-0.476). <br>
We continue with the next variable

**BsmtFinSF2: Type 2 finished square feet**  

```{r}
plot(data$BsmtFinSF2,data$SalePrice)
cor(data[1:1459,'SalePrice'],data[1:1459,'BsmtFinSF2'])
data$BsmtFinSF2[is.na(data$BsmtFinSF2)]
paste(colnames(select(data,contains("Bsmt"))),collapse="','")
data[is.na(data$BsmtFinSF2),c('BsmtQual','BsmtCond','BsmtExposure','BsmtFinType1','BsmtFinSF1','BsmtFinType2','BsmtFinSF2','BsmtUnfSF','TotalBsmtSF','BsmtFullBath','BsmtHalfBath')]
data$BsmtFinSF2[is.na(data$BsmtFinSF2)]<-0 # Replace missing value with zero
cor(data$BsmtFinSF1,data$BsmtFinSF2)
```

We will probably drop this variable,let's see if it is correlated with another

```{r}
cor(data$BsmtFinSF2,data[,sapply(data, is.numeric)])
```

Correlated with BsmtFinType2 <br>
Now we see the correlations between basement variables

```{r}
paste(colnames(select(data,contains("Bsmt"))),collapse="','")
cor(data[,c('BsmtQual','BsmtCond','BsmtExposure','BsmtFinType1','BsmtFinSF1','BsmtFinType2','BsmtFinSF2','BsmtUnfSF','TotalBsmtSF','BsmtFullBath','BsmtHalfBath')])
```

We will create a corrplot for better understanding the correlations

```{r}
corrplot(cor(data[,c('BsmtQual','BsmtCond','BsmtExposure','BsmtFinType1','BsmtFinSF1','BsmtFinType2','BsmtFinSF2','BsmtUnfSF','TotalBsmtSF','BsmtFullBath','BsmtHalfBath')])
,method = "square")
```

The basement quality and basement condition (which are important variables) are not correlated with BsmtHalfBath and BsmtFinSF2. So we will drop these two variables

```{r}
data <- subset(data, select = -BsmtFinSF2)
data <- subset(data, select = -BsmtHalfBath)
```


##*Now that we have taken care of Na's let's see which are the numeric variables.*##


```{r}
colnames(data[,sapply(data, is.numeric)]) #check out which columns are numeric 
```

Let's see also some correlations. <br>
We will find the numeric variables that have the highest correlation with the Saleprice variable and examine them further.

```{r}
corrplot(cor(na.omit(data[,sapply(data, is.numeric)])),method = "square")
paste(which(cor(data[1:1459,sapply(data, is.numeric)],data[1:1459,'SalePrice'])>0.5 | cor(data[1:1459,sapply(data, is.numeric)],data[1:1459,'SalePrice'])<(-0.5)),collapse=',') # Check for high or low correlations
```

colnames(data[,sapply(data,is.numeric)][c(5,7,8,11,17,20,23,27,30,31,32,44)])


**Id**
We will get rid of the Id column and keep it in a vector called Id

```{r}
Id<-data$Id[1461:nrow(data)]
data <- data[,-1]
```

**MSSubClass: Identifies the type of dwelling involved in the sale.**

```{r}
plot(data$MSSubClass,data$SalePrice)
```

In reality this is a factor and not numeric. 

```{r}
table(as.factor(data$MSSubClass))
data$MSSubClass<-as.factor(data$MSSubClass) # Change into factor variable
```

**OverallQual: Rates the overall material and finish of the house**

```{r}
table(as.factor(data$OverallQual))
plot(as.factor(data$OverallQual),data$SalePrice)
data$OverallQual<-as.integer(data$OverallQual) # Change into integer variable
table(data$OverallQual)
```

**OverallCond: Rates the overall condition of the house**

```{r}
table(as.factor(data$OverallCond))
plot(as.factor(data$OverallCond),data$SalePrice)
data$OverallCond<-as.integer(data$OverallCond)
table(data$OverallCond)
```

**YearBuilt: Original construction date**

```{r}
plot(as.factor(data$YearBuilt))
table(as.factor(data$YearBuilt))
plot(as.factor(data$YearBuilt),data$SalePrice)
cor(data[1:1459,'SalePrice'],data[1:1459,'YearBuilt'])
```

**YearRemodAdd: Remodel date (same as construction date if no remodeling or additions)**

```{r}
plot(as.factor(data$YearRemodAdd))
table(as.factor(data$YearRemodAdd))
plot(as.factor(data$YearRemodAdd),data$SalePrice)
plot(data$YearRemodAdd,data$YearBuilt)
#We can see from the plot that all remodelings started on 1950. Is it true or there is some error?
cor(data$YearRemodAdd>1950,data$YearBuilt>1950) #Highly correlated with Yearbuilt
```

**Pool Area**

```{r}
table(data$PoolArea) # We'll drop this variable (most observations belong to one parameter)
data <- subset(data, select = -c(PoolArea))
```

Afer examining the above numeric variables, we also need to examine the remaining character variables.

```{r}
colnames(data[,sapply(data, is.character)]) 
```
 

**PoolQC**

```{r}
table(data$PoolQC) #we'll drop this variable
data <- subset(data, select = -c(PoolQC))
```



**Street: Type of road access to property **

```{r}
plot(as.factor(data$Street))
table(data$Street)
#Only 12 Gravels      
#removing Street variable 
data <- subset(data, select = -Street)
```



